---
title: "Appendix 1: A Detailed R Notebook With Code"
output:
  html_notebook: default
  pdf_document: default
---
A simple R script was submitted for Turnitin. This is a clearer notebook which allows for more readability, but cannot be copy/pasted by the marker to check the code
# Read the dataset
```{r}
#setwd("C:/Users/Ariel/Google Drive/UNI/ACTL2102/Assignment")
claims = read.csv('claimsdata.csv')
attach(claims)
head(claims)
```

#Create the transition matrix

The table function creates a contingency table of the counts of each variable. In other words, the number of people that transition from state i to j (Nij). Dividing these numbers by the number of people in state i (Ni*) gives the estimates for transition probability.

```{r}
P = table(claims$ncdlevel18, claims$ncdlevel19)
Ni = rowSums(P)
P = (P/Ni)
P
```

# Conduct a simulation to estimate 2020 premium income using 2019 state

The simplest way to do this would be to define a function. The arguments of the function are the transition probability matrix, the initial state of the insured life, and the number of simulations to run. The function is defined as follows:

  1. Create an empty vector 'sim.j' with the same length as the number of simulations to run. This will house our result from each simulation.
  2. The 2019 Discount Level determines initial state i and thus the row of the Transition Probability Matrix to use. This is given by the argument 'init', and we add three to this number for appropriate indexing
  3. Create a loop which will use a randomly generated multinomial (with probability distribution equal to the i'th row of the transition matrix) to determine the state j (in 2020). Add this to the vector 'sim.j'
  4. Use the function that outputs premium from discount level to determine the 2020 premium income
  5. Output the mean of all simulated premiums for that row

```{r}
Run_MC_Sim <- function(PT.matrix = P, init = 0, num.sims = 1000){
  sim.j = numeric(num.sims)
  # init + 3 --> To match to index
  p = PT.matrix[init+3, ]
  for (n in 1:num.sims){
    sim.j[n] = which(rmultinom(1,1,p) == 1) - 3
  }
  Premium = 300*(1-(sim.j*0.1))
  return(mean(Premium))
}
```

Finally, we run our function on the NCDlevel19 variable to produce the simulated expected premium income. The sum of these is our expected total premium income.

```{r}
claims$Simulated.Premium = rep(0,length(claims$X))

for (i in 1:length(claims$X) ){
  claims$Simulated.Premium[i]  = Run_MC_Sim(PT.matrix = P, init = claims$ncdlevel19[i], num.sims = 1000)
}

sum(claims$Simulated.Premium)
```

# 2018 Profitability
```{r}
claims$Premium2018 = 300*(1-(claims$ncdlevel18*0.1))
claims$ClaimsCost2018 = 2000*pmin(claims$numberofclaims , 2)
LossRatio = sum(claims$ClaimsCost2018)/sum(claims$Premium2018)
LossRatio
```


# Poisson Model
```{r}
Lambda = mean(numberofclaims)
Lambda
```

```{r}
Run_Poisson_Sim <- function(Lambda, init = 0, num.sims = 1000){
  sim.j = numeric(num.sims)
  for (n in 1:num.sims){
    # Number of accidents is poisson RV, with roof at 2
    accidents = max(rpois(1,Lambda),2)
    # If no accidents, move up a level
    if (accidents == 0){state.j = init + 1
    # if 1, 2, drop 1,2 levels. level floor at -2
    } else {state.j = max(init - accidents, -2)}
    sim.j[n] = state.j
  }
  Premium = 300*(1-(sim.j*0.1))
  return(mean(Premium))
}
```

```{r}
claims$Simulated.Premium.Poisson = rep(0,length(claims$X))

for (i in 1:length(claims$X) ){
  claims$Simulated.Premium.Poisson[i]  = Run_Poisson_Sim(Lambda = Lambda, init = claims$ncdlevel19[i], num.sims = 1000)
}

head(claims)
```

```{r}
sum(claims$Simulated.Premium.Poisson) - sum(claims$Simulated.Premium) 
```

